RED WINE ANALYSIS by Trang Nguyen

========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

#install.packages("ggplot2")
#install.packages('gridExtra')
#install.packages('GGally')
#install.packages('corrplot')
library(corrplot)
library(ggplot2)
library(gridExtra)
library(GGally)
library(plyr)
library(dplyr)
library(cowplot)
library(RColorBrewer)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
#getwd()
setwd('C:/Users/nguyentt/Documents/Courses/DAND/P4')
wine <- read.csv('wineQualityReds.csv')
#names(wine)

grey_pallete <- c("#f7f7f7","#d9d9d9","#bdbdbd","#969696","#636363","#252525") #for quality
blue_pallete <- c("#eff3ff","#bdd7e7","#6baed6","#2171b5") #for alcohol
green_pallete <- c("#edf8e9","#c7e9c0","#a1d99b","#74c476","#31a354","#006d2c") #for acid citric

#create new variables
wine$quality.factor <- as.factor(wine$quality)
wine$citric.acid.bracket <- cut(wine$citric.acid,breaks = seq(-0.01,1,0.15))
wine$alcohol.bracket = cut(wine$alcohol,breaks=c(8,10,12,14,16))

#univariate common plotting function

ggplot_univ_common <- function(feature, bins){
  ggplot(data = wine, aes_string(x = feature)) + geom_histogram(binwidth = bins)
}

##bivariate common scatter plotting
ggplot_bi_common <- function(x_feature, y_feature, x_label, y_label, chart_title){
  ggplot(aes_string(x=x_feature,y=y_feature), data=wine) + 
  geom_point(color = "#969696") +
  geom_smooth(color = "black", se = FALSE, method = "lm")+ 
  labs(title = chart_title, y = y_label, x = x_label)
}

ggplot_density_common<- function(x_feature,color_factor,color_palette,
                                 legend_name,x_label,y_label,chart_title){
  ggplot(wine, aes_string(x_feature, color = color_factor)) +
  geom_line(stat="density",size = 1.5)+
  scale_color_manual(values = color_palette, name = legend_name)+theme_dark()+ 
  labs(title = chart_title, y = y_label, x = x_label, legend = legend_name)
}

##multivariate common scatter plotting function

ggplot_point_multi_common <- function(x_feature,y_feature,color_factor,
                                dot_size,alpha_param,legend_name,x_label,
                                y_label,chart_title){
  ggplot(data=wine, aes_string(x=x_feature,y=y_feature,color = color_factor))+
    geom_point(size=dot_size,alpha = alpha_param) + theme_dark()+
    labs(title = chart_title, y = y_label, x = x_label, legend = legend_name)
}

ggplot_jitter_multi_common <- function(x_feature,y_feature,color_factor,
                                       color_palette,dot_size,alpha_param,
                                       legend_name,x_label,y_label,chart_title){
  ggplot(data=wine, aes_string(x=x_feature,y=y_feature,color = color_factor))+
    geom_jitter(size=dot_size,alpha = alpha_param) + theme_dark()+
    labs(title = chart_title, y = y_label, x = x_label, legend = legend_name)+
    scale_color_manual(values = color_palette, name = legend_name)
}


ggplot_density_scatter_combine <- function(x_feature,y_feature,color_factor,
                                           color_palette,dot_size,
                                           alpha_param,x_label,y_label,
                                           chart_title){
  
  p <-ggplot_point_multi_common(x_feature,y_feature,color_factor,dot_size,
                                           alpha_param,"","","","")+
    scale_color_manual(values = green_pallete, name = "Quality")  +
    geom_smooth(se = FALSE, method = "lm") +theme_dark()
  
  p2<-p+ theme(legend.position="none",axis.title.x=element_blank(),
             axis.title.y=element_blank())+labs(title = "")
  
  p1 <- ggplot_density_common(x_feature,color_factor,color_palette,
                              "",x_label,y_label, chart_title)+
    theme(legend.position="none", axis.title.y=element_blank())
  
  p3<-ggplot_density_common(y_feature,color_factor,color_palette,
                              "",y_label,"", "")+
    theme_dark()+ coord_flip() + theme(legend.position="none",
                                     axis.title.x=element_blank())
  
  legend <- g_legend(p)
  grid.arrange(arrangeGrob(p1,legend, p2, p3, nrow=2))
}

g_legend<-function(a.gplot){
    tmp <- ggplot_gtable(ggplot_build(a.gplot))
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
    legend <- tmp$grobs[[leg]]
    return(legend)
}

```

# Univariate Plots Section

In this section, I would explore the distributions of all variables.

Below is a summary of all variables in the dataset.

```{r echo=FALSE,warning=FALSE}

summary(wine)
```

Next, I'm going to investigate the distribution of each variable.

####Fixed Acidity plot

```{r echo=FALSE,warning=FALSE}
qplot(x=fixed.acidity,data=wine,binwidth = .1, xlab='Fixed Acidity, g/dm^3', 
      main = "Fixed Acidity Histogram") + scale_x_continuous(breaks=4:16)

summary(wine$fixed.acidity)
```

####Volatile Acidity plot

```{r echo=FALSE,warning=FALSE}
qplot(x=volatile.acidity,data=wine,binwidth = 0.01,
      xlab='Volatile Acidity, g/dm^3', main = "Volatile Acidity Histogram") + 
  scale_x_continuous(breaks=pretty(wine$volatile.acidity,n=20))

summary(wine$volatile.acidity)
```

####Citric Acid plot

There is quite a significant number of wines with 0 amount of citric acid:  132 wines (~8% of the samples). There are also some peaks (more than 50 counts) at ranges (0.02,0.03], (0.2,0.21], (0.23, 0.24], (0.48,0.49].

```{r echo=FALSE,warning=FALSE}

qplot(x=citric.acid,data=wine,binwidth = 0.01, xlab = 'Citric Acid, g/dm^3',
      main = "Citric Acid Histogram")

cut_citric.acid <- cut(wine$citric.acid,breaks = seq(-0.01,1,0.01))

table(cut_citric.acid)

summary(wine$citric.acid)

```

####Residual Sugar plots

Since the original histogram for residual sugar is right skewed, I log-transformed it.

```{r echo=FALSE,warning=FALSE}

qplot(x=residual.sugar,data=wine,binwidth=0.2, xlab = 'Residual Sugar, g/dm^3',
      main = "Residual Sugar Histogram")+
  scale_x_continuous(breaks = 0:16)

ggplot(aes(x=residual.sugar),data=wine)+
  scale_x_log10() + xlab('Residual Sugar, log transformed') + 
  labs(title = "Resiudal Sugar Histogram, Log Transformed")+
  geom_histogram(bins = 60)

summary(wine$residual.sugar)

```

####Chlorides plot

There are a big number of ouliers for chlorides. After limit to 3% and 97% quantiles for this feature, I log transformed to get a more normalized distribution.

```{r echo=FALSE,warning=FALSE}

qplot(x=chlorides,data=wine,binwidth=0.01, xlab = "Chlorides, g/dm^3", 
      main = "Chlorides Histogram") + 
  scale_x_continuous(breaks = seq(0,0.6,0.05))

ggplot_univ_common("chlorides",0.002)+
  xlab('Chlorides from 0.03 quantile to 0.97 quantile, g/dm^3')+
  xlim(quantile(wine$chlorides,0.03),quantile(wine$chlorides,0.97))+
  labs(title = "Chlorides Histogram, Remove Outliers")

ggplot_univ_common("chlorides",0.025)+
  xlab('Chlorides from 0.03 quantile to 0.97 quantile, log transformed, g/dm^3')+
  xlim(quantile(wine$chlorides,0.03),quantile(wine$chlorides,0.97))+
  scale_x_log10()+
  labs(title = "Chlorides Histogram, Remove Outliers, Log Transformed")

summary(wine$chlorides)
```

####Free Sulfur Dioxide plot

Based on the table, we can see that there is a big number of wines which have free sulfur dioxide of 5 and 6 mg/dm^3. 104 wines have free sulfur dioxide of 5 and 138 wines of 6. The histogram for free sulfur dioxide is right skewed so I log transformed it.

```{r echo=FALSE,warning=FALSE}

qplot(x=free.sulfur.dioxide,data=wine,binwidth=1,
      xlab='Free Sulfur Dioxide, mg/dm^3', 
      main = "Free Sulfur Dioxide Histogram")

ggplot_univ_common("free.sulfur.dioxide",0.05)+
  xlab('Free Sulfur Dioxide, log transformed, mg/dm^3')+ scale_x_log10() + 
  labs(title = "Free Sulfur Dioxide Histogram, Log Transformed")

table(wine$free.sulfur.dioxide)
summary(wine$free.sulfur.dioxide)

```

####Total Sulfur Dioxide plot

Based on the table below, we can see that there are more a third of wines which have total sulfur dioxide from 15 to 30 mg/dm^3. That is a significant number given that the max total sulfur dioxide is 289 mg/dm^3. The histogram for total sulfur dioxide is right skewed so I log transformed it.

```{r echo=FALSE,warning=FALSE}

qplot(x=total.sulfur.dioxide,data=wine,binwidth=5,
      xlab = 'Total Sulfur Dioxide, mg/dm^3', 
      main = "Total Sulfur Dioxide Histogram")

ggplot_univ_common("total.sulfur.dioxide",0.03)+
  xlab('Total Sulfur Dioxide, log transformed, mg/dm^3')+
  scale_x_log10() + 
  labs(title = "Total Sulfur Dioxide Histogram, Log Transformed")

cut_total.sulfur.dioxide <- cut(wine$total.sulfur.dioxide,
                                breaks = seq(0,300,5))
table(cut_total.sulfur.dioxide)

summary(wine$total.sulfur.dioxide)
```

####Density plot

```{r echo=FALSE,warning=FALSE}

qplot(x=density,data=wine,binwidth=0.00025, xlab = 'Density, g/cm^3',
      main = "Density Histogram")

summary(wine$density)
```

####pH plot

```{r echo=FALSE,warning=FALSE}

qplot(x=pH,data=wine,binwidth=0.05, xlab = 'pH', main = "pH Histogram")
summary(wine$pH)
```

####Sulphates plot

I log transformed sulphates to get a more normalized distribution.

```{r echo=FALSE,warning=FALSE}

qplot(x=sulphates,data=wine,binwidth=0.025, xlab = 'Sulphates, g/dm3',
      main = "Sulphates Histogram")

ggplot_univ_common("sulphates",0.025)+ scale_x_log10() + 
  xlab('Sulphates, log transformed, g/dm3') + 
  labs(title = "Sulphates Histogram, Log Transformed")

summary(wine$sulphates)
```

####Alcohol plot

Based on the table, we can see that there are more wines which have alcohol percentage of from 9.2 to 10 (686 wines ~ 43% of total samples).

```{r echo=FALSE,warning=FALSE}

qplot(x=alcohol,data=wine,binwidth=0.25, xlab = 'Alcohol, percent by vol',
      main = "Alcohol Histogram")

summary(wine$alcohol)
table(wine$alcohol)

table(wine$alcohol.bracket)
```

####Quality plot

The majority of the wines have quality 5 and 6 (82% of all samples). 13.6% of all samples have quality 7 and 8.

```{r echo=FALSE,warning=FALSE}

qplot(x=quality.factor,data=wine,xlab = 'Quality', main = "Quality Histogram")

table(wine$quality.factor)

```

# Univariate Analysis

#### What is the structure of your dataset?
This dataset consists of rating and 11 attributes of 1599 types of wine. The attributes are:

1. fixed.acidity
2. volatile.acidity
3. citric.acid
4. residual.sugar     
5. chlorides
6. free.sulfur.dioxide
7. total.sulfur.dioxide
8. density
9. pH
10. sulphates
11. alcohol

Quality has values from 3 to 8 where wine types having quality 5 and 6 accounting for more than 80% of the total.

#### What is/are the main feature(s) of interest in your dataset?

With this analysis, I would like to find which attributes are most likely to effect the quality of the wine. I also did some reading online to have a background for this study. The main features of interest that I think I will focus on more are alcohol, citric acid and residual sugar.

#### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

I think the combination of fixed and volatile acidity as well as density can support the investigation also.

#### Did you create any new variables from existing variables in the dataset?

I created brackets for alcohol and citric acid to help with the analysis later. For alcohol, I have 4 brackets: 8-10%, 10-12%, 12-14% and 14-16%. For citric acid, I have 6 brackets: 0-0.14, 0.14-0.29, 0.29-0.44, 0.44-0.59, 0.59-0.74 and 0.74-0.89 g/dm^3.

#### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
For some of the positive skewed attributes, I had to transform them into log histogram.


# Bivariate Plots Section

Firstly, I created ggcorr plot to get the overall relationships between the variables in the dataset.

```{r echo=FALSE, Bivariate_Plots,warning=FALSE}

##ggpair plots
exclude <- c("X", "quality.factor", "citric.acid.bracket","alcohol.bracket")
keep <- !(names(wine) %in% exclude)
ggcorr(wine[,keep],nbreaks = 5, label = TRUE, label_size = 3,
       color = "grey50", size = 2) + 
  labs(title = "GGCorr graph for all features in dataset")
```

We can see some relationships between the variables as below:

Strong positive correlation (>0.6):

- Fixed Acidity and Citric Acid
- Fixed Acidity and Density
- Free Sulfure Dioxide and Total Sulfur Dioxide

Strong negative correlation (<-0.6):

- Fixed Acidity and pH

Medium positive correlation (>0.2, <0.6):

- Volatile Acidity and pH
- Citric Acid and Chlorides
- Citric Acid and Density
- Citric Acid and Sulphates
- Citric Acid and Quality
- Residual Sugar and Total Sulfur Dioxide
- Residual Sugar and Density
- Chlorides and Density
- Chlorides and Sulphates
- pH and Alcohol
- Sulphates and Quality
- Alcohol and Quality

Medium negative correlation (>-0.5, <-0.3):

- Fixed Acidity and Volatile Acidity
- VOlatile Acidity and Citric Acid
- VOlatile Acidity and Sulphates
- VOlatile Acidity and Alcohol
- Volatile Acidity and Quality
- Citric Acid and pH
- Chlorides and pH
- Chlorides and Alochol
- Total Sulfur Dioxide and Alcohol
- Density and Alcohol
- Density and pH

```{r echo=FALSE,warning=FALSE}

```
####Alcohol vs Quality

As we can see from the box plot, higher quality wines have higher percentage of alcohol. However, there are not many wines that have very high percentage of alcohol; so it does not guarantee that higher percentage of alcohol will make the wine taste good.

```{r echo=FALSE,warning=FALSE}
qplot(x=quality.factor,y=alcohol,data=wine,geom='boxplot',xlab = "Quality", 
      ylab = "Alcohol, percent by volume", main = "Boxplot for Quality and Alcohol")

cor.test(wine$quality,wine$alcohol)

ggplot_density_common("alcohol","quality.factor",green_pallete,"Quality",
                      "Alcohol, percent by volume","Density, g/dm3",
                      "Alcohol Density Graph, by Quality")

tapply(wine$alcohol,wine$quality.factor,summary)

```

####Volatile Acidity vs Quality

As we can see from the charts, wines with lower amount of volatile acidity seem to have higher quanlity points.

```{r echo=FALSE,warning=FALSE}
qplot(x=quality.factor,y=volatile.acidity,data=wine,geom='boxplot',
      xlab = "Quality", ylab = "Volatile Acidity , g/dm^3",
      main = "Box Plot for Volatile Acidity and Quality")

cor.test(wine$quality,wine$volatile.acidity)

ggplot_density_common("volatile.acidity","quality.factor",green_pallete,"Quality",
                      "Volatile Acidity , g/dm^3","Density, g/dm3",
                      "Volatile Acidity Density Graph, by Quality")

tapply(wine$volatile.acidity,wine$quality.factor,summary)

```

Below are scatter plots for other pairs of variables with strong correlations that we found.

#### Fixed Acidity vs Citric Acid

Fixed Acidity and Citric Acid has strong positive correlation (0.67 from Pearson's test) as in the graph.

```{r echo=FALSE,warning=FALSE}

ggplot_bi_common("fixed.acidity","citric.acid","Fixed Acidity ,g/dm^3","Citric Acid, g/dm^3",
                 'Fixed Acidity vs Citric Acid scatter plot') +
  xlim(quantile(wine$fixed.acidity,0.01), quantile(wine$fixed.acidity,0.99)) +
  ylim(quantile(wine$citric.acid,0.01), quantile(wine$citric.acid, 0.99))
```

#### Fixed Acidity vs Density

Fixed Acidity and pH has strong positive correlation (0.67 from Pearson's test)

```{r echo=FALSE,warning=FALSE}

ggplot_bi_common("fixed.acidity","density","Fixed Acidity ,g/dm^3","Density, g/dm3",
                 "Fixed Acidity vs Density scatter plot")

```

#### Fixed Acidity vs pH

Fixed Acidity and pH has strong negative correlation (-0.68 from Pearson's test)

```{r echo=FALSE,warning=FALSE}

ggplot_bi_common("fixed.acidity","pH","Fixed Acidity ,g/dm^3" ,"pH",
                 'Fixed Acidity vs pH scatter plot')

```

#### Free Sulfur Dioxide vs Total Sulfur Dioxide

Free Sulfur Dioxide and Total Sulfur Dioxide unsurprisingly has high positive correlation (0.67 from Pearons's test).

```{r echo=FALSE,warning=FALSE}

ggplot_bi_common("free.sulfur.dioxide","total.sulfur.dioxide",
                 "Free Sulfur Dioxide, mg / dm^3","Total Sulfur Dioxide, mg / dm^3",
                 'Free Sulfur Dioxide vs Total Sulfur Dioxide scatter plot')+
  xlim(quantile(wine$free.sulfur.dioxide,0.01),
       quantile(wine$free.sulfur.dioxide,0.99)) +
  ylim(quantile(wine$total.sulfur.dioxide,0.01),
       quantile(wine$total.sulfur.dioxide, 0.99))

```

#### Citric Acid vs Volatile Acidity

Citric Acid and Volatile Acidity are 0.55 correlated (from Pearson's test).

```{r echo=FALSE,warning=FALSE}

ggplot_bi_common("citric.acid","volatile.acidity","Citric Acid, g/dm^3" ,
                 "Volatile Acidity , g/dm^3",
                 'Citric Acid vs Volatile Acidity scatter plot')

```

#### Citric Acid vs pH

Citric Acid and pH has 0.54 Pearson's correlation.

```{r echo=FALSE,warning=FALSE}


ggplot_bi_common("citric.acid","pH","Citric Acid, g/dm^3" ,"pH",
                 'Citric Acid vs pH scatter plot')


```


# Bivariate Analysis

#### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

The investigations about relationships between features were based on Pearson correlation.

Citric acid has:
- large positive correlation with fixed acidity (0.67),
- large negative correlation with volatile acidity (-0.55),
- large negative correlation with pH (-0.54),
- medium postivie correlation with density (0.36)

Residual sugar has:
- medium positive correlation with density (0.36)
- 
Alcohol has:
- medium positive correlation with quality (0.48)
- medium negative correlation with density (-0.5).


#### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

Fixed acidity has:
- large positive correlation with density (0.67)
- large negative correlation with pH (-0.68)

Volatile acidity has medium negative correlation with quality (-0.39)

Chlorides has medium positive correlation with sulphates (0.37)

Free sulfur has strong postivie correlation with total sulfur dioxide (0.67)

Density has
- medium negative correlation with ph
- medium negative correlation alcohol

#### What was the strongest relationship you found?
The strongest relationship is between fixed acidity and pH (0.68). This makes sense because pH is defined as the measurement for how acidic or basic a wine is on a scale of 0 (very acidic) to 14 (very basic). For that reason, the more fixed acidity in the wine, the higher pH becomes.


# Multivariate Plots Section


####Citric Acid, Fixed Acidity and Quality

This is a combined plot of a scatter plot and two density plots for fixed acidity and citric acid with the colors of the points are by quality points.It seems like higher quality wines have higher fixed acidity and higher citric acid.

```{r echo=FALSE, Multivariate_Plots,warning=FALSE}

ggplot_density_scatter_combine("citric.acid","fixed.acidity","quality.factor",
                               green_pallete,2, 0.5,"Citric Acid, g/dm^3",
                               "Fixed Acidity ,g/dm^3",
                               "Citric Acid vs Fixed Acidity, by Quality")

```



####Citric Acid, Volatile Acidity and Quality

I have another combined plot of a scatter plot and two density plots for volatile acidity and citric acid with the colors of the points are by quality points. In this plot, the higher quality wines seem to concentrate where citric acid is high and volatile acidity is low.

```{r echo=FALSE,warning=FALSE}

ggplot_density_scatter_combine("citric.acid","volatile.acidity","quality.factor",
                               green_pallete,2, 0.5,"Citric Acid, g/dm^3",
                               "Volatile Acidity , g/dm^3",
                               "Citric Acid vs Volatile Acidity, by Quality")
```



####Citric Acid, pH and Quality

In this plot, higher quality wines have higher citric acid and lower pH.

```{r echo=FALSE,warning=FALSE}

ggplot_density_scatter_combine("citric.acid","pH","quality.factor",
                               green_pallete,2,0.5,"Citric Acid, g/dm^3","pH",
                               "Citric Acid vs pH, by Quality")

```



####Citric Acid, Density and Quality

This combined plot shows the relationship between citric acid and quality. Density doesn't seem to interact with quality here.

```{r echo=FALSE,warning=FALSE}

ggplot_density_scatter_combine("citric.acid","density","quality.factor",
                               green_pallete,2,0.5,"Citric Acid, g/dm^3","Density, g/dm3",
                               "Citric Acid vs Density, by Quality")

```



####Citric Acid, Sulphates and Quality

With the density plot of sulphates factored by quality, we can see that wines with higher sulphates and higher citric acid have better quality.

```{r echo=FALSE,warning=FALSE}

ggplot_density_scatter_combine("citric.acid","sulphates","quality.factor",
                               green_pallete,2,0.5,"Citric Acid, g/dm^3","Sulphates, g /dm^3",
                               "Citric Acid vs Sulphates, by Quality")
```



####Density, Alcohol and Quality

It is not very clear; however we can see more wines with high quality on the upper left portion of the plot: higher alcohol percentage and smaller density.

```{r echo=FALSE,warning=FALSE}

ggplot_density_scatter_combine("density","alcohol","quality.factor",
                               green_pallete,2,0.5,"Density, g/dm3","Alcohol, percent by volume",
                               "Density vs Alcohol, by Quality")
```



####Volatile Acidity, Quality and Alcohol Brackets

This is a different kind of scatter plot. We have volatile acidity vs quality and the points are colored by alcohol brackets. As we can see, darker colored dots have higher quality and lower volatile acidity.

```{r echo=FALSE,warning=FALSE}
ggplot_jitter_multi_common("volatile.acidity","quality","alcohol.bracket",
                           blue_pallete,2,1,
                           "Alcohol, percent by volume","Volatile Acidity , g/dm^3","Quality",
                           "Volatile Acidity vs Quality, by Alcohol")

```



####Density, Quality and Alcohol Brackets

With the plot below, we can not really see a clear relationship between the three variables.

```{r echo=FALSE,warning=FALSE}

ggplot_jitter_multi_common("density","quality","alcohol.bracket",blue_pallete,
                           2,1,"Alcohol, percent by volume","Density, g/dm3","Quality",
                           "Density vs Quality, by Alcohol")
  
```



####pH, Quality and Citric Acid Brackets.

This plot has pH vs quality and the points are colored by citric acid brackets. As we can see, there are more darker colored dots at the upper left of the plot which means that higher quality wines tend to have higher citric acid and lower pH.

```{r echo=FALSE,warning=FALSE}
ggplot_jitter_multi_common("pH","quality","citric.acid.bracket",green_pallete,2,1,
                           "Citric Acid, g/dm^3","pH","Quality",
                           "pH vs Quality, by Citric Acid")+
  scale_x_continuous(breaks = seq(0,4,0.1))

summary(subset(wine,wine$citric.acid==0))

```

# Multivariate Analysis

#### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

For a lot of cases, citric acid when considered with other variables gives stronger indication for higher quality wines. We can see that the combinations of citric acid and fixed acidity, volatile acidity and pH really strengthen each other in accordance to our interest of quality values. 

#### Were there any interesting or surprising interactions between features?

I think it is interesting that there are a lot of wines which have citric acid amount equals to zero but still it seems to be an important feature in the analysis.

#### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

#### Plot One
```{r echo=FALSE, Plot_One,warning=FALSE}
ggplot_density_common("alcohol","quality.factor",green_pallete,"Quality",
                      "Alcohol, percent by volume","Density, g/dm3","Alcohol, percent by volume")
```

#### Description One

This plot shows the density of alcohol percentage by volume factored by quality scores. The reason why I chose it was becasue from the chart, we can see cleary the positive correlation between alchol and quality. Most of the higher quality wines are distributed towards higher alcohol direction.

#### Plot Two
```{r echo=FALSE, Plot_Two,warning=FALSE}

ggplot_density_scatter_combine("citric.acid","volatile.acidity","quality.factor",
                               green_pallete,2, 0.5,"Citric Acid, g/dm^3",
                               "Volatile Acidity , g/dm^3",
                               "Citric Acid vs Volatile Acidity, by Quality")
```

#### Description Two

I think this plot is good in showing the relationship between Volatile Acidity, Citric Acid and Quality. We can see that wines with higher citric acid and lower volatile acidity received better quality scores.

#### Plot Three
```{r echo=FALSE, Plot_Three,warning=FALSE}
ggplot_jitter_multi_common("volatile.acidity","quality","alcohol.bracket",
                           blue_pallete,2,1,
                           "Alcohol, percent by volume","Volatile Acidity , g/dm^3","Quality",
                           "Volatile Acidity vs Quality, by Alcohol")
```

#### Description Three

I like this plot because we see the darker color dots (for higher alcohol) are distributed mostly at the upper left corner of the chart where we have higher quality score and lower volatile acidity. In conclusion, higher alcohol and lower density seem to produce higher quality wines.

------

# Reflection

This project is really helpful for students who are learning to practice exploratory data analysis. R is a very powerful tool to assist us with our tasks of univariate, bivariate and multivariate plotting and analysys. In bivariate analysys, we found the correlations between pairs of variables in the dataset. However, there might be interactions between more than two variables, that's why multivariate analysis was needed in the project. 

I learnt that choosing the right colors in the plots really helps with the process. I used the default palette initially but was advised to use a sequential palette from RColorBrewer which was really useful in identifying trends and relationships. I used GGpairs for getting the overall idea on relationships between variables. After the first project review, I was advised to use GGCorr which really made it easier for me to visualize the correlations.

There are other combinations of variables that haven't been investigated in this study such as residual and density or chlorides and sulphates (both have medium correlation). Besides, different types of charts on the same variables can also reveal interesting relationships that we haven't found. Future approaches can study this dataset from those different angles. And since some variables seem to have linear correlation with quality, we can apply a regression model to build a machine learning model for wine quality prediction.
